\newcommand{\thesispointone}{A methodology and case study for code synthesis \\ \mbox{evaluation} of LLMs}
\newcommand{\thesispointtwo}{A program synthesis dataset for LLM temperature \\ \mbox{analysis}}
\newcommand{\thesispointthree}{Assessment of GPT-4 in Fixing Real-World Software \\ \mbox{Vulnerabilities}}

\newcommand{\thesispointoneuf}{A methodology and case study for code synthesis evaluation of LLMs}
\newcommand{\thesispointtwouf}{A program synthesis dataset for LLM temperature analysis}
\newcommand{\thesispointthreeuf}{Assessment of GPT-4 in Fixing Real-World Software Vulnerabilities}

\newcommand{\thesispointonehu}{Módszertan és esettanulmány a nagy nyelvi modellek kódszintézis-értékeléséhez}
\newcommand{\thesispointtwohu}{Programszintézis adathalmaz a nagy nyelvi modellek temperature elemzéséhez}
\newcommand{\thesispointthreehu}{A GPT-4 értékelése valós szoftversérülékenységek \\ javításában}

\newcommand{\thesisonetext}{I reviewed existing LLM comparison methodologies and identified their limitations, which led me to design a more comprehensive evaluation approach.
I selected a suitable benchmark, evaluated multiple language models on it, and performed a quality analysis of the synthesized programs.
In addition, I designed and implemented a human evaluation framework to assess developer acceptance of LLM-generated code and demonstrated the methodology through a case study.}

\newcommand{\thesistwotext}{I designed the study to create the dataset containing 18,900 raw and 18,896 processed C++ code generation outputs from large language models.
I developed a model selection framework to systematically identify suitable open-source models and evaluated nine LLMs from three model families with parameter sizes up to 70 billion using varying temperature settings.
Finally, I analyzed the results to investigate the effect of the temperature hyperparameter on source code generation and showed that it does not reveal a clear performance pattern.}


\newcommand{\thesisthreetext}{I explored prompt engineering techniques to generate fixed source code with GPT-4 for automatic program repair, aiming to maintain functionality while removing vulnerabilities.
I designed the evaluation framework for both prompts and models, using the Vul4J benchmark and an additional post-training benchmark to ensure unbiased assessment.
I also participated in the manual evaluation of the generated source code and textual responses, finding that GPT-4 could fix 33.33\% of vulnerabilities and provide useful guidance in roughly 50\% of cases.}