\renewcommand{\chapterautorefname}{Chapter}

In the \textbf{first thesis group}, the contributions are related to Large Language Model comparison in software synthesis. Detailed discussion can be found in \autoref{chapter_2}.

\begin{enumerate}[wide = 0pt, widest = {I/4.}, leftmargin =*]
	
	\item[I/1.] I collected relating methodologies and determined the gaps in those methodologies.
	
	\item[I/2.] I created the comparison methodology.
	
	\item[I/3.] I searched for an evaluation benchmark.
	
	\item[I/4.] I evaluated the Language Models on the benchmark.
	
	\item[I/5.] I performed the quality analysis of the synthetized programs.
	
	\item[I/6.] I designed the Human Evaluation.
	
	\item[I/7.] I created the Human Evaluation framework.
	
\end{enumerate}

\vspace{1cm}

\noindent
In the \textbf{second thesis group}, the contributions are related to creation of a dataset and the hyper parameter investigation of Large Language Models in software synthesis. Detailed discussion can be found in \autoref{chapter_4}.

\begin{enumerate}[wide = 0pt, widest = {II/5.}, leftmargin =*]
	
	\item[II/1.] I designed the study.
	
	\item[II/2.] I created the model selection framework.
	
	\item[II/3.] I created the model evaluation framework.
	
	\item[II/4.] I evaluated the data according to the temperature variable effect.
	
\end{enumerate}

\vspace{1cm}

\noindent
In the \textbf{third thesis group}, the contributions are related to the evaluation of GPT-4 in real-life vulnerability fixing. Detailed discussion can be found in \autoref{chapter_3}.

\begin{enumerate}[wide = 0pt, widest = {III/5.}, leftmargin =*]
	
	\item[III/1.] I searched for prompt engineering techniques.
	
	\item[III/2.] I designed the evaluation of the prompts and also the models.
	
	\item[III/3.] I participated in the manual evaluation of the generated source code.
	
	\item[III/4.] I participated in the manual evaluation of the generated textual responses.
	
\end{enumerate}

The author states that while the thesis results are primarily his own work, the
pronoun we is used instead of I to recognize the input of co-authors in the papers
forming the basis of this thesis.
